---
title: "Assignment 2 PSYR6003"
output: html_document
date: "2025-03-14"
---

#First we are setting up our librarys for our analysis then reading in our data from our repository using read_sav.
```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(flexplot)
library(haven)
library(apaTables)
library(psych)
df <- read_sav('PSYR6003.A2.sav')
```

##Cleaning and Setting up Analysis
#To set up the analysis we must convert sex from a string to binary. In addition one participant has "Other" as their sex, we have removed them from the analysis due to being the only participant and not being specified otherwise. 
#We can then use dyplr to change strings of 'Male' to 1, and everything else becomes 0.
```{R}
df <- df %>% filter(sex != 'Other (specify)')
df <- df %>%
  mutate(sex = ifelse(sex == 'Male', 1, 0))
#Use table to see that it worked
table(df$sex)
```

#Reversing then taking mean of each subscale
#We can use the case_when function to say which number should be converted into another number. This allows us to specifically reverse the order. 

#Following this we can use the rowMeans for each variable we want to add up, I simply type every specificaly subscale together and put it into a new variable with .TOT in it.

#After that we can use "select" and head to see if it worked.
```{R}
table(df$tipm.CONS2.3y)
df <- df %>%
  mutate(across(tipm.CONS2.3y, 
                ~ case_when(
                  . == 1 ~ 7,
                  . == 2 ~ 6,
                  . == 3 ~ 5,
                  . == 5 ~ 3,
                  . == 6 ~ 2,
                  . == 7 ~ 1
                )))

#After this we can 
df$tipm.TOT <- rowMeans(df[c("tipm.CONS1.3y", "tipm.CONS2.3y")], na.rm = TRUE)

df$panas.TOT <- rowMeans(df[c("guilt1.3y", "guilt2.3y", "guilt3.3y", 
                              "dep1.3y", "dep2.3y", "dep3.3y", 
                              "fear1.3y", "fear2.3y", "fear3.3y", 
                              "host1.3y", "host2.3y", "host3.3y")], na.rm = TRUE)

df$mps.TOT <- rowMeans(df[c("mps.SPP1.3y", "mps.SPP2.3y", "mps.SPP3.3y", 
                            "mps.SPP4.3y", "mps.SPP5.3y")], na.rm = TRUE)
select <- df %>% 
  select(c('tipm.TOT', 'panas.TOT', 'mps.TOT', 'sex'))
head(view, 20)
```

##Utilizing flexplot we can look at the distribution of each of our variables.
#Based on a visual look, none of the data looks very normally distributed or equally distributed in the case of male vs female.

```{R}
#Examine our data firstly
#Predictors
flexplot(sex ~ 1, data = df)
flexplot(tipm.TOT ~ 1, data = df)
flexplot(mps.TOT~  1, data = df)
#Outcome
flexplot(panas.TOT ~ 1 , data = df)
```
##Examine for Assumptions, first we set up our model then use the visualize function to visualize what the residuals of our model look like.
#This model gives us somewhat normally distributed residuals and variance that isnt constant. 
```{r}
model <- lm(panas.TOT~mps.TOT + tipm.TOT + sex,
data=df)
visualize(model, plot="residuals")
visualize(model)
```
##Examine our model now by creating APA tables of descriptive statistics and a table for our model
#Use summary model to examine our model firstly.
#Following this we can plug it into apa.reg.table in order to get an APA table to put into our paper. 
#I have removed manually things from this output that were unnecessary for our results, in addition because it did not have the SE I manually added it into the table. 
```{R}
summary(model)
apa.reg.table(model, filename = "Regression_Table.doc")
```


#To create a descriptive APA table, we can use the "select" dataframe from before, which only includes sex/tipm/panas/mps, and create a new dataframe that will get the mean and SD. 
#apa.cor.table will then make a table with correlation between the two coefficients as well. 
```{r}
descriptives <- data.frame(
  Variable = colnames(select),
  Mean = sapply(select, mean, na.rm = TRUE),
  SD = sapply(select, sd, na.rm = TRUE)
)
apa.cor.table(select, filename = "correlation_table.doc", table.number = 1)
```

##Finally model comparison
#We recreate the full model from before
#Then we create a reduced model that takes away the main variable of interest (mps.TOT)
#Following that we can use the model.comparison function to compare our models in a variety of way, seeing that the full is better on all accounts.
```{r}

full <- lm(panas.TOT~sex + tipm.TOT + mps.TOT,
data=df)
reduced <- lm(panas.TOT~sex + tipm.TOT,
data=df)
model.comparison(reduced, full)
```